---
title: '"Find and Replace" in Multiple Variable Names'
slug: "find-replace-in-var-names"
date: 2017-10-04
comments: true
draft: true
categories: ["R"]
tags: ["R", "tidyverse", "dplyr", "purrr"]

showpagemeta: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(blogdown)
library(emo)
```


Renaming a variable/set of variables or column names is fairly straightforward. And there are plenty of resources on The Google. If you're familiar with the `dplyr` package in R, you've probably used `select()` and `rename()` a lot. What may not be as straight forward to a beginner or intermediate R user is how to rename a group of variables at the same time or "find and replace" a string of text in a group of variable names---as opposed to making the changes one by one.

This issue has come up most frequently for me in 3 ways:

1. I download a dataset and a group of variables has a simlar "prefix" or "suffix" added to the variable name. E.g. `prefix_varname1`, `prefix_varname2` or `varname1_suffix`, `varname2_suffix`.

2. I "collapse" or "summarise" whole or part of a dataset based on some common transformation, like a mean. And I get a new variable/column name like `varname1_mean`, `varname2_mean` as an end product.

3. I join/merge two or more large datasets and there are identical variable names in those datasets, and I see something like `var1.x` or `var1.y`

Using just `select()` or `rename()` you can change these variable names one by one to another name. You might do something like this...


```{r the-problem, message=FALSE, warning=FALSE}
library(tidyverse)
library(stringr)

## CREATE THE FAKE DATA
df <- tribble(
  ~prefix_x, ~prefix_y, ~prefix_z, # column names
#-----------|---------|---------|
  "country1",   100,    1,         # values in each column
  "country2",   500,    2
)

df

## RENAME YOUR VARIABLES
df %>%
  rename(x = prefix_x,
         y = prefix_y,
         z = prefix_z
         )
```


This is a perfectly good solution. Nothing wrong with using simple tools! This gets time consuming though if you have a lot of variables.

If you're a beginner, or even intermediate, you may not have read the `dplyr` documentation thoroughly because there's so much to consume already `r emo::ji("anguished")`. But, if you glanced at it you might have noticed a reference to the "scoped" variants of `select()` or `rename()`--- namely, `select_all`, `select_if`, `select_at`, `rename_all`, `rename_if`, `rename_at`. In R, type `?select` and you'll see what I'm talking about. These functions are designed to tackle our problem, but I remember there being a bit of a hurdle to understand how to actually use them.

There are **many** ways to solve this problem, and I've included a few examples of how to do this because one way might be easier for you to remember than another. I repeat...there are many!



### Solution 1

Initially, I found this example to be easiest to remember because it didn't require knowledge of something like "quosure" or `purrr`. First, we'll create the data frame. Second, we'll "find and replace"/delete a string of text in the variable name using a `stringr` package function `str_replace_all`. This package and it's many functions are the `tidyverse` way to use regex [^1] for string/text manipulation.

[^1]: "regular expressions"

```{r solution1, warning=FALSE, message=FALSE}
## FAKE DATA
df2 <- tribble(
  ~x_mean, ~y_mean, ~z_mean,  # column names
  #------|---------|-------|
   2.5,    100,     1,        # values in each column
   5,      500,     0.5
)

df2

## RENAME THE VARS BY TAKING OUT "_mean"
df2 %>%
  dplyr::rename_all(
    funs(stringr::str_replace_all(., "_mean", ""))
    )
```

I'll explain really quickly what a few parts of the code do because it might help you remember how to write it in the future. `.` in `str_replace_all()` is a place holder of sorts for the data frame object, `df2`, we already created. The next argument in the function is the pattern we want to find in the variable name, i.e. `"_mean"`. The last argument, is what we want to replace it with. In this case we just want to "delete" the variable names' suffix.

`funs()` is one way to call an outside tool/function in the "scoped" `dplyr` functions, and add some expression inside of that called tool/function.



### Solution 2
```{r solution2, warning=FALSE, message=FALSE}
## RENAME THE VARS BY TAKING OUT "_mean"
df2 %>%
  dplyr::rename_all(
    ~stringr::str_replace_all(., "_mean", "")
    )
```
This example uses purrr-style notation. You can read more about it in the `purrr` documentation.[^2]


[^2]: useful sites for reading more if unfamiliar: [RStudio's purrr intro](http://purrr.tidyverse.org) and ["jennybc"'s purrr tutorial](https://jennybc.github.io/purrr-tutorial/)


### Solution 3
```{r solution3, warning=FALSE, message=FALSE}
library(purrr)
# or... `library(tidyverse)` if you loaded that already

## RENAME THE VARS BY TAKING OUT "_mean"
df2 %>%
  set_names(~stringr::str_replace_all(., "_mean", "")
            )
```
This uses the `set_names` function in `purrr`.

Like I mentioned above, there are a lot of different tools out there to do this. Instead of using the `stringr` package and functions, you might like using base R's `grep` functions. Those would work here too by just substituting them for `stringr::str_replace_all`.
